---
title: "T-cell dynamics"
author: "GM + ?"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Using mixed models to fit T-cell dynamic models

Subjects are observed at a number of occasions at different times. On each occasion the densities of cells of five different types are recorded.

The process for these five types, $N, S, C, T$ and $E$, are modelled as follows:

\[\begin{aligned}
  {N_{t + 1}} =  & {N_t} + {B_t} - N_t^d - N_t^t \\ 
  {S_{t + 1}} =  & {S_t} + N_t^t - S_t^d - S_t^t \\ 
  {C_{t + 1}} =  & {C_t} + S_t^t - C_t^d - C_t^t \\ 
  {T_{t + 1}} =  & {T_t} + C_t^t - T_t^d - T_t^t \\ 
  {E_{t + 1}} =  & {E_t} + T_t^t - E_t^d \\ 
\end{aligned} \]

where 

1. $X_t^d$ represents the number of cells of type $X$ that die between 
time $t$ and time $t+1$,
2. $X_t^t$ represents the number of cells of type $X$ that transition to the next type,
3. $B_t$ represents the number of new cells of type $N$.

Letting $d_t = \tau_{t+1} - \tau_t$ be the number of units of time elapsed between the the observation at time $\tau_t$ and that at time $\tau_{t+1}$, the 
hypothetical distribution of each component is: 

\[\begin{aligned}
  {B_t}\sim & Poisson(\lambda d_t) \\
  N_t^d\sim & Poisson\left( {{N_t}{\delta _N d_t}} \right)\\
  N_t^t\sim & Poisson\left( {{N_t}{\psi _N} d_t} \right)\\
  S_t^d\sim & Poisson\left( {{S_t}{\delta _S} d_t} \right) \\
  S_t^t\sim & Poisson\left( {{S_t}{\psi _S} d_t} \right)  \\
  C_t^d\sim & Poisson\left( {{C_t}{\delta _C} d_t} \right) \\
  C_t^t\sim & Poisson\left( {{C_t}{\psi _C} d_t} \right) \\
  T_t^d\sim & Poisson\left( {{T_t}{\delta _T} d_t} \right)\\
  T_t^t\sim & Poisson\left( {{C_t}{\psi _T} d_t} \right) \\
  E_t^d\sim & Poisson\left( {{E_t}{\delta _E} d_t} \right)  \\
\end{aligned} \]

where $\lambda$ represents the expected number of new cells of type $N$ per unit of time and the 9 parameters of the form $\delta_X, \psi_X$, represent 
the probability of death or transition per cell per unit of time. If $d_t$
is close to constant, it can be be omitted in the model.

This model should be accurate if the number of dying and transitioning cells of
each type is relatively small compared with the total number of cells of 
that type within each observed time period. In this case, it may be reasonable to assume that the $B_t, N_t^d, N_t^t, ...$ components are independent within occasion $t$ and sequentially conditionally independent, i.e. the distribution of $B_t |N_t$ is independent of the distribution of $B_{t+1}|N_{t+1}$, etc.

The change in cell counts between occasions is:

\[\Delta \left[ {\begin{array}{*{20}{c}}
  {{N_{t + 1}}} \\ 
  {{S_{t + 1}}} \\ 
  {{C_{t + 1}}} \\ 
  {{T_{t + 1}}} \\ 
  {{E_{t + 1}}} 
\end{array}} \right] = \left[ {\begin{array}{*{20}{c}}
  1&{ - 1}&{ - 1}&0&0&0&0&0&0&0 \\ 
  0&0&1&{ - 1}&{ - 1}&0&0&0&0&0 \\ 
  0&0&0&0&1&{ - 1}&{ - 1}&0&0&0 \\ 
  0&0&0&0&0&0&1&{ - 1}&{ - 1}&0 \\ 
  0&0&0&0&0&0&0&0&1&{ - 1} 
\end{array}} \right]\left[ {\begin{array}{*{20}{c}}
  {{B_t}} \\ 
  {N_t^d} \\ 
  {N_t^t} \\ 
  {S_t^d} \\ 
  {S_t^t} \\ 
  {C_t^d} \\ 
  {C_t^t} \\ 
  {T_t^d} \\ 
  {T_t^t} \\ 
  {E_t^d} 
\end{array}} \right]\]

\[\operatorname{Var} \left( {\Delta \left[ {\begin{array}{*{20}{c}}
  {{N_{t + 1}}} \\ 
  {{S_{t + 1}}} \\ 
  {{C_{t + 1}}} \\ 
  {{T_{t + 1}}} \\ 
  {{E_{t + 1}}} 
\end{array}} \right]} \right) = A\left[ {diag\left( {\begin{array}{*{20}{c}}
  \lambda &{{\delta _N}{N_t}}&{{\psi _N}{N_t}}&{{\delta _S}{S_t}}&{{\psi _S}{S_t}}&{{\delta _C}{C_t}}&{{\psi _C}{C_t}}&{{\delta _T}{T_t}}&{{\psi _T}{T_t}}&{{\delta _E}{E_t}} 
\end{array}} \right)} \right]A'\]

\[A = \left[ {\begin{array}{*{20}{c}}
  1&{ - 1}&{ - 1}&0&0&0&0&0&0&0 \\ 
  0&0&1&{ - 1}&{ - 1}&0&0&0&0&0 \\ 
  0&0&0&0&1&{ - 1}&{ - 1}&0&0&0 \\ 
  0&0&0&0&0&0&1&{ - 1}&{ - 1}&0 \\ 
  0&0&0&0&0&0&0&0&1&{ - 1} 
\end{array}} \right]\]

\[\left[ {\begin{array}{*{20}{c}}
  {\lambda  + {N_t}\left( {{\delta _N} + {\psi _N}} \right)}&{ - {\psi _N}{N_t}}&0&0&0 \\ 
  { - {\psi _N}{N_t}}&{{\psi _N}{N_t} + {S_t}\left( {{\delta _S} + {\psi _S}} \right)}&{ - {\psi _S}{S_t}}&0&0 \\ 
  0&{ - {\psi _S}{S_t}}&{{\psi _S}{S_t} + {C_t}\left( {{\delta _C} + {\psi _C}} \right)}&{ - {\psi _C}{C_t}}&0 \\ 
  0&0&{ - {\psi _C}{C_t}}&{{\psi _C}{C_t} + {T_t}\left( {{\delta _T} + {\psi _T}} \right)}&{ - {\psi _T}{T_t}} \\ 
  0&0&0&{ - {\psi _T}{T_t}}&{{\psi _T}{T_t} + {\delta _E}{E_t}} 
\end{array}} \right]\]

\[\left[ {\begin{array}{*{20}{c}}
  {\lambda  + {N_t}{\phi _N}}&{ - {\psi _N}{N_t}}&0&0&0 \\ 
  { - {\psi _N}{N_t}}&{{\psi _N}{N_t} + {S_t}{\phi _S}}&{ - {\psi _S}{S_t}}&0&0 \\ 
  0&{ - {\psi _S}{S_t}}&{{\psi _S}{S_t} + {C_t}{\phi _C}}&{ - {\psi _C}{C_t}}&0 \\ 
  0&0&{ - {\psi _C}{C_t}}&{{\psi _C}{C_t} + {T_t}{\phi _T}}&{ - {\psi _T}{T_t}} \\ 
  0&0&0&{ - {\psi _T}{T_t}}&{{\psi _T}{T_t} + {E_t}{\phi _E}} 
\end{array}} \right]\]

\[E\left( {\Delta \left[ {\begin{array}{*{20}{c}}
  {{N_{t + 1}}} \\ 
  {{S_{t + 1}}} \\ 
  {{C_{t + 1}}} \\ 
  {{T_{t + 1}}} \\ 
  {{E_{t + 1}}} 
\end{array}} \right]} \right) = \left[ {\begin{array}{*{20}{c}}
  {\lambda  - {\delta _N}{N_t} - {\psi _N}{N_t}} \\ 
  {{\psi _N}{N_t} - {\delta _S}S - {\psi _S}{S_t}} \\ 
  {{\psi _S}{S_t} - {\delta _C}{C_t} - {\psi _C}{C_t}} \\ 
  {{\psi _C}{C_t} - {\delta _T}{T_t} - {\psi _T}{T_t}} \\ 
  {{\psi _T}{T_t} - {\delta _E}{E_t}} 
\end{array}} \right] = \left[ {\begin{array}{*{20}{c}}
  {\lambda  - {N_t}\left( {{\delta _N} + {\psi _N}} \right)} \\ 
  {{\psi _N}{N_t} - {S_t}\left( {{\delta _S} + {\psi _S}} \right)} \\ 
  {{\psi _S}{S_t} - {C_t}\left( {{\delta _C} + {\psi _C}} \right)} \\ 
  {{\psi _C}{C_t} - {T_t}\left( {{\delta _T} + {\psi _T}} \right)} \\ 
  {{\psi _T}{T_t} - {\delta _E}{E_t}} 
\end{array}} \right] = \left[ {\begin{array}{*{20}{c}}
  {\lambda  - {N_t}{\phi _N}} \\ 
  {{\psi _N}{N_t} - {S_t}{\phi _S}} \\ 
  {{\psi _S}{S_t} - {C_t}{\phi _C}} \\ 
  {{\psi _C}{C_t} - {T_t}{\phi _T}} \\ 
  {{\psi _T}{T_t} - {E_t}{\phi _E}} 
\end{array}} \right]\]


== Notes on vignettes ==
Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` setion of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
